Specific leaf mass explains difference in biomass partitioning between woody plant functional types

### 2014-10-15 manuscript/draftpaper.Rnw
intro + methods

\item Major plant functional types differ widely in specific leaf mass (SLM, the ratio of leaf mass to area, an index of leaf thickness) (\cite{poorter_causes_2009}), which can be expected to affect biomass partitioning in several ways. First, SLM correlates with leaf lifespan. Longer lived leaves would lead to higher plant leaf mass (MF) as long as annual production rates are similar (\cite{enquist_global_2002}). Another interpretation is that plants may target a certain leaf area display given a certain biomass investment in supporting tissues, because light interception and photosynthesis correlate with the exchange area of the plant.

\item It has been reported that gymnosperms (with higher SLM) have higher leaf mass fraction than angiosperms (with lower SLM) \cite{enquist_global_2002, mccarthy_consistency_2007, poorter_biomass_2012, sardans_tree_2013}. But a recent study reported no difference (\cite{wolf_allometric_2010}). No study to date has interpreted differences in partitioning between PFTs in terms of SLM. This would be useful because SLM is already an important parameter in DGVMs, potentially reducing parameters required for the allocation subroutine.

\item What is lacking is a solid empirical base that can be used to constrain any model of allocation, that includes woody plants of all sizes, of all major plant functional types and biomes. Most global efforts to date use the Cannell database of forest biomass, which is not based on individual biomass estimates. Here we use a new database (Falster et al.), the largest ever compiled for individual woody plants, to evaluate whether important differences exist in biomass partitioning woody plant functional types. We also test whether these differences are confounded (or compounded) by climate.

@@

@@

>>>>

\item Major plant functional types differ widely in specific leaf mass (SLM, the ratio of leaf mass to area, an index of leaf thickness) (\cite{poorter_causes_2009}), which can be expected to affect biomass partitioning in several ways. First, SLM correlates with leaf lifespan. Longer lived leaves would lead to higher plant leaf mass ($M_F$) as long as annual production rates are similar (\cite{enquist_global_2002}). Another interpretation is that plants may target a certain leaf area display given a certain biomass investment in supporting tissues, because light interception and photosynthesis correlate with the exchange area of the plant.

\item It has been reported numerous times that gymnosperms (with higher SLM) have higher leaf mass fraction than angiosperms (with lower SLM) \cite{enquist_global_2002, mccarthy_consistency_2007, poorter_biomass_2012, sardans_tree_2013}, although a recent study reported no difference (\cite{wolf_allometric_2010}). No study to date has been able to analyze differences between deciduous and evergreen angiosperms, or link differences in partitioning between PFTs to other plant traits such as SLM.

\item What is lacking is a solid empirical base that can be used to constrain any model of allocation, that includes woody plants of all sizes, of all major plant functional types and biomes. Most global efforts to date use the Cannell database of forest biomass, which is not based on individual biomass estimates. Here we use a new database (Falster et al.), the largest ever compiled for individual woody plants, including data for major plant functional types (PFTs), and all major biomes.

\item We consider here the relationship between leaf biomass and area with stem diameter (an important allometric relationship, Enquist others, and related to the pipe model if the ratio sapwood / heartwood is relatively constant, an important measure of hydraulic supply), the aboveground leaf mass fraction (partioning of biomass into leaf vs. woody pools), and the leaf area ratio (a measure of the photosynthetic area per unit woody biomass). We expect that $M_F$ and $A_F$ are approximately proportional to the square of stem diameter, which would remove size-related shifts in allometry that often confound interpretations. For each of these variables, we ask whether 1) there are differences between major plant functional types after removing size-related variation, 2) there are differences due to climate, compared between major biomes, or as a function of the mean annual temperature (as recently reported for whole stand biomass, \cite{reich_temperature_2014}).

@@

\item For the aboveground biomass pools, we calculated the leaf mass fraction ($M_F / M_W$) and leaf area ratio ($A_F / M_W$). LMF, LAR, SLM, $M_F$ and $A_F$ are related by,

$$LMF = SLM * LAR = (M_F / A_F) * LAR$$

@@

\item A recent study compared allocation routines in a number of leading ecosystem models \cite{de2014does}, and recommended to constrain allocation by observed biomass fractions instead of using constant allocation fractions, which is common practice.
